Modeling browser-based distributed evolutionary computation systems

### 2015-02-08 modeling-volunteer-computing-gecco.tex
Solving conflict

In this paper we have presented our experience on using browser-based
computing applied
to the design of evolutionary algorithms.

There are many issues involved in using these resources: from adapting current algorithms so that they match this environment \cite{agajaj} to check which EA configuration works the best in it, or creating a framework that can use it easily \cite{nodeo2014}, or even security issues. But the main challenge is that asking people to contribute resources implies that you are opening your science to society and you have to give something in return: you have to adopt a set of best practices that have come to be known as Open Science, because ``Give, and it shall be given unto you'', you will get as much back from society as you give to it opening your science and explaining it to the public. This, among other things, means that popularity will become directly performance of the metacomputer you create by attracting more users.

>>>>

In this paper we have presented our experience on using browser-based
computing applied to the design of evolutionary algorithms. In the
spirit of Open Science, we have released all material related to the
experiment, including this paper. Experiments have been performed at
several times of the day and announced in Twitter or during a live
presentation about distributed computation using the browser.

What we conclude from these computers is that the nature of the
performance, which is due to the number of persons deciding to
participate in the experiment, is not totally random. There are at
least lower bounds we can count on: several computers (as few as six)
can almost always be relied on, and in some cases up to 30 can
participate in a single experiment. However, it is not clear how all
these computers contribute to the overall performance in terms of
time, although all experiments were performed until a solution was
found. Most computers participating in the experiment are in the same
performance tier, although a small percentage of them (which will be
around 25\%) are quite slow.

The evolutionary algorithm programmed in this way can easily
accommodate all kinds of browsers. However, we are using an homogeneous
configuration for all of them, which might not be the best from an
algorithmic point of view; a random parameter setting like the one
used by \cite{LNCS86720702} could offer better results, since we do
not know in advance what is the performance of the nodes. More
systematic experimentation is also needed, specially using different
kind of problems, including more complex problems in which the fitness
function is {\em heavier} and takes longer to be processed. However,
the main intention of this paper, which was to prove that distributed
evolutionary computation could be done efficiently in a volunteer/free
environment, has been sufficiently proved.

There are many issues involved in using these resources: from adapting
current algorithms so that they match this environment 
to check which EA configuration works the best in it, or creating a
framework that can use it easily. But the main
challenge is that asking people to contribute resources implies that
you are opening your science to society and you have to give something
in return: you have to adopt a set of best practices that have come to
be known as Open Science, because ``Give, and it shall be given unto
you'', you will get as much back from society as you give to it
opening your science and explaining it to the public. This, among
other things, means that popularity will become directly performance
of the metacomputer you create by attracting more users. And this
interplay between the social network itself (popularity in Twitter,
people attending to the conference and interested in following the
slides in their computer) is another, very interesting, venue to
explore in the future.

\section{Acknowledgements}

This paper is supported in part by several grants hidden for
anonymity.
